### 2022-07-15 chapters/interpretability/imli/iterative_learning.tex
considering all minor comments

In iterative learning, we learn one clause of a CNF classifier in each iteration, where the clause refers to a partial classifier. The coverage of a clause in a CNF formula is the set of samples that \emph{do not satisfy} the clause. The reason is that if a sample does not satisfy at least one clause in a CNF formula, the prediction of the sample by the full formula is class $ 0 $ because CNF is a conjunction of clauses. As a result, considering covered samples in the next iteration does not change their prediction regardless of whatever clause we learn in later iterations. To this end, a single clause learning can be performed efficiently by applying mini-batch learning discussed before. In Algorithm~\ref{interpretability_imli_algo:iterative_CNF_learning}, we provide an algorithm for learning a CNF classifier iteratively by leveraging mini-batch learning. This algorithm is a double-loop algorithm, where in the outer loop we apply iterative learning and in the inner loop, we apply mini-batch learning.

@@

\paragraph{Terminating conditions.} In Algorithm~\ref{interpretability_imli_algo:iterative_CNF_learning}, we terminate iterative learning based on three conditions: (i) when $ \Rule $ contains all $ k $ clauses , (ii) the training data $ \dataset $ is empty (that is, no sample remains uncovered), and (iii) no new sample is covered by the partial classifier. Since the first two conditions are trivial, we elaborate on the third condition. When clause $ C_i $ cannot cover any new sample from the training dataset $ \dataset $, the next iteration will result in the same clause $ C_i $ because the training data remains the same. In this case, we do not include clause $ C_i $ to classifier $ \Rule $ because of zero coverage.

@@

\subsection{Learning DNF classifiers} 
\label{interpretability_imli_sec:dnf_learning}
For learning DNF classifiers, we leverage De Morgan's law where complementing a CNF formula results in a DNF formula. To learn a DNF classifier, say $ \Rule'(\nonsensitive) $, we can trivially show that $ Y = \Rule'(\nonsensitive) \leftrightarrow \neg (Y = \neg \Rule'(\nonsensitive)) $ for the feature vector $ \nonsensitive $. Here $ \neg \Rule'(\nonsensitive) $ is a CNF formula, by definition. Thus, we learn a DNF classifier by first complementing the class-label $ y $ to $ \neg y $ for each sample in the training dataset $ \{(\mathbf{x}^{(i)}, y^{(i)})\}_{i=1}^n $, learning a CNF classifier on $ \{(\mathbf{x}^{(i)}, \neg y^{(i)})\}_{i=1}^n $ and finally complementing the learned classifier to DNF. For example, the CNF classifier  ``(Male $ \vee $ Age $ < 50 $) $ \wedge $ (Education $ = $ Graduate $ \vee $ Income  $ \ge 1500 $)'' 	is complemented to a DNF classifier as 	 ``(not Male $ \wedge $ Age $ \ge 50 $) $ \vee $ (Education $ \ne $ Graduate $ \wedge $ Income $ \le 1500 $)''.

@@

\begin{algorithm}
	\caption{Iterative learning of decision lists}
	\label{interpretability_imli_algo:iterative_decision_lists_learning}
	\begin{algorithmic}[1]
		\Procedure{DecisionListLearning}{$ \dataset, \lambda, k $}
		\State $ \mathcal{R}_L \leftarrow \{\}$
		\For{$ i \leftarrow 1, \dots, k-1 $ and $ \dataset \ne \emptyset $}
		\State $ V_i \leftarrow \textsc{MajorityClass}(\{y^{(i)}\}_{i=1}^{|\dataset|}) $ \Comment{$ V_i $ specifies the target class given class label}
		\State $ C_i \leftarrow \textsc{MiniBatchDNFLearning}( \dataset, \lambda, 1, V_i) $ \Comment{Ref. Section~\ref{interpretability_imli_sec:dnf_learning}}

@@

\begin{algorithm}
	\caption{Iterative learning of decision sets}
	\label{interpretability_imli_algo:iterative_decision_sets_learning}
	\begin{algorithmic}[1]
		\Procedure{DecisionSetsLearning}{$\dataset,\lambda, k$}
		\State $ \mathcal{R}_S=\{\} $
		\State $ \dataset_{cc} = \{\}$  \Comment{Contains correctly covered samples}
		\For{$ i \leftarrow 1, \dots, k-1 $ and $ \dataset \ne \emptyset $}
		\State $ V_i \leftarrow \textsc{MajorityClass}(\{y^{(i)}\}_{i=1}^{|\dataset|}) $
		\State $ \dataset_{w} = \dataset  \cup \{(\mathbf{x}, \neg V_i) | (\mathbf{x}, y) \in \dataset_{cc}\} $ \Comment{Appending covered samples with complemented class }
%		\State $ \mathbf{X}^\text{w} \leftarrow \mathbf{X} \cup \mathbf{X}^\text{cc} $ \Comment{Correctly covered samples are included}
%		\State $ \mathbf{y}^\text{w} \leftarrow  \mathbf{y} \cup  \{\neg V_i\}^{|\mathbf{X}^\text{cc}|}$ \Comment{$ \mathbf{X}^\text{cc} $ have complemented class $ \neg V_i $ to minimize overlap}
		\State $ C_i \leftarrow \textsc{MiniBatchDNFLearning}( \dataset_w,\lambda, 1, V_i) $ 
		\State $ \dataset' \leftarrow \textsc{CorrectCoverage}(\dataset, C_i, V_i) $
		\If{$ (\dataset' = \emptyset $} 
		\State break
		\EndIf
		\State $ \mathcal{R}_S \leftarrow \mathcal{R}_S \cup \{(C_i, V_i)\} $
		\State $ \dataset \leftarrow \dataset  \setminus \dataset' $
		\State $ \dataset_{cc} = \dataset_{cc} \cup \dataset' $
%		\State $ \mathbf{X}^\text{cc} = \mathbf{X}^\text{cc} \cup \mathbf{X}'$ 
		\EndFor
		\State $ \mathcal{R}_S \leftarrow \mathcal{R}_S  \cup \{(\text{true}, V_{\mathsf{default}} )\}$
		\State \Return $ \Rule_S $
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

\begin{itemize}
	\item The first modification is with respect to the definition of coverage for decision sets. Unlike decision lists, we separate samples that are \textit{correctly covered} by $ (C_i, V_i) $ in each iteration. Given  $ (\mathbf{X}, \mathbf{y}) $, the correctly covered samples of $ (C_i, V_i) $ is a dataset $ \dataset_{cc}  =  \{(\mathbf{x}^{(l)}, y^{(l)}) | \mathbf{x}^{(l)} \models C_i \wedge y^{(l)} = V_i \}_{l=1}^{|\dataset|} \subseteq \dataset $ of samples that satisfy $ C_i $ and have matching class-label as $ V_i $. 
	\item The second modification is related to the  training dataset considered in each iteration. Let $ \dataset $ denote \emph{the remaining training dataset} in the current iteration and $ V_i $ be the majority class in  $ \dataset $. Hence, $ V_i $ is the target class in the current iteration. Also, let $ \dataset_{cc} $ denotes the set of samples \emph{correctly covered in all previous iterations}. In Algorithm~\ref{interpretability_imli_algo:iterative_decision_sets_learning}, we learn the current rule $ C_i $ on the working dataset $ \dataset_{w} = \dataset  \cup \{(\mathbf{x}, \neg V_i) | (\mathbf{x}, y) \in \dataset_{cc}\} $. Thus, $ \dataset_{w} $ is constructed by joining the remaining training samples with correctly covered samples in $ \dataset_{cc} $ with complemented class label to the target class $ V_i $.  By explicitly labeling covered samples as class $ \neg V_i $, the new clause $ C_i $ learns to falsify already covered samples. This heuristic allows us to minimize the overlap of $ C_i $ compared to previously learned clauses $ \{C_j\}_{j=1}^{i-1} $.
\end{itemize}

>>>>

In iterative learning, we learn one clause of a CNF classifier in each iteration, where the clause refers to a partial classifier. The coverage of a clause in a CNF formula is the set of samples that \emph{do not satisfy} the clause. The reason is that if a sample does not satisfy at least one clause in a CNF formula, the prediction of the sample by the full formula is class $ 0 $, because CNF is a conjunction of clauses. As a result, considering covered samples in the next iteration does not change their prediction regardless of whatever clause we learn in later iterations. To this end, a single clause learning can be performed efficiently by applying mini-batch learning discussed before. In Algorithm~\ref{interpretability_imli_algo:iterative_CNF_learning}, we provide an algorithm for learning a CNF classifier iteratively by leveraging mini-batch learning. This algorithm is a double-loop algorithm, where in the outer loop we apply iterative learning and in the inner loop, we apply mini-batch learning.

@@

\paragraph{Terminating conditions.} In Algorithm~\ref{interpretability_imli_algo:iterative_CNF_learning}, we terminate iterative learning based on three conditions: (i) when $ \Rule $ contains all $ k $ clauses , (ii) the training data $ \dataset $ is empty (that is, no sample remains uncovered), and (iii) no new sample is covered by the current partial classifier. Since the first two conditions are trivial, we elaborate on the third condition. When clause $ C_i $ cannot cover any new sample from the training dataset $ \dataset $, the next iteration will result in the same clause $ C_i $ because the training data remains the same. In this case, we do not include clause $ C_i $ to classifier $ \Rule $ because of zero coverage.

@@

\subsection{Learning DNF classifiers} 
\label{interpretability_imli_sec:dnf_learning}
For learning DNF classifiers, we leverage De Morgan's law where complementing a CNF formula results in a DNF formula. To learn a DNF classifier, say $ \Rule'(\nonsensitive) $, we can trivially show that $ Y = \Rule'(\nonsensitive) \leftrightarrow \neg (Y = \neg \Rule'(\nonsensitive)) $ for the feature vector $ \nonsensitive $. Here $ \neg \Rule'(\nonsensitive) $ is a CNF formula, by definition. Thus, we learn a DNF classifier by first complementing the class-label $ y^{(i)} $ to $ \neg y^{(i)} $ for each sample in the training dataset $ \{(\mathbf{x}^{(i)}, y^{(i)})\}_{i=1}^n $, learning a CNF classifier on $ \{(\mathbf{x}^{(i)}, \neg y^{(i)})\}_{i=1}^n $, and finally complementing the learned classifier to DNF. For example, the CNF classifier  ``(Male $ \vee $ Age $ < 50 $) $ \wedge $ (Education $ = $ Graduate $ \vee $ Income  $ \ge 1500 $)'' 	is complemented to a DNF classifier as 	 ``(not Male $ \wedge $ Age $ \ge 50 $) $ \vee $ (Education $ \ne $ Graduate $ \wedge $ Income $ \le 1500 $)''.

@@

\begin{algorithm}
	\caption{Iterative Learning of Decision Lists}
	\label{interpretability_imli_algo:iterative_decision_lists_learning}
	\begin{algorithmic}[1]
		\Procedure{DecisionListLearning}{$ \dataset, \lambda, k $}
		\State $ \mathcal{R}_L \leftarrow \{\}$
		\For{$ i \leftarrow 1, \dots, k-1 $ and $ \dataset \ne \emptyset $}
		\State $ V_i \leftarrow \textsc{MajorityClass}(\{y^{(i)}\}_{i=1}^{|\dataset|}) $ \Comment{$ V_i $ specifies the target class given class label}
		\State $ C_i \leftarrow \textsc{MiniBatchDNFLearning}( \dataset, \lambda, 1, V_i) $ \Comment{Ref. Section~\ref{interpretability_imli_sec:dnf_learning}}

@@

\begin{algorithm}
	\caption{Iterative Learning of Decision Sets}
	\label{interpretability_imli_algo:iterative_decision_sets_learning}
	\begin{algorithmic}[1]
		\Procedure{DecisionSetsLearning}{$\dataset,\lambda, k$}
		\State $ \mathcal{R}_S=\{\} $
		\State $ \dataset_{cc} = \{\}$  \Comment{Contains correctly covered samples}
		\For{$ i \leftarrow 1, \dots, k-1 $ and $ \dataset \ne \emptyset $}
		\State $ V_i \leftarrow \textsc{MajorityClass}(\{y^{(i)}\}_{i=1}^{|\dataset|}) $
		\State $ \dataset_{w} = \dataset  \cup \{(\mathbf{x}, \neg V_i) | (\mathbf{x}, y) \in \dataset_{cc}\} $ \Comment{Appending covered samples with complemented class }
%		\State $ \mathbf{X}^\text{w} \leftarrow \mathbf{X} \cup \mathbf{X}^\text{cc} $ \Comment{Correctly covered samples are included}
%		\State $ \mathbf{y}^\text{w} \leftarrow  \mathbf{y} \cup  \{\neg V_i\}^{|\mathbf{X}^\text{cc}|}$ \Comment{$ \mathbf{X}^\text{cc} $ have complemented class $ \neg V_i $ to minimize overlap}
		\State $ C_i \leftarrow \textsc{MiniBatchDNFLearning}( \dataset_w,\lambda, 1, V_i) $ 
		\State $ \dataset' \leftarrow \textsc{CorrectCoverage}(\dataset, C_i, V_i) $
		\If{$ (\dataset' = \emptyset $} 
		\State break
		\EndIf
		\State $ \mathcal{R}_S \leftarrow \mathcal{R}_S \cup \{(C_i, V_i)\} $
		\State $ \dataset \leftarrow \dataset  \setminus \dataset' $
		\State $ \dataset_{cc} = \dataset_{cc} \cup \dataset' $
%		\State $ \mathbf{X}^\text{cc} = \mathbf{X}^\text{cc} \cup \mathbf{X}'$ 
		\EndFor
		\State $ \mathcal{R}_S \leftarrow \mathcal{R}_S  \cup \{(\text{true}, V_{\mathsf{default}} )\}$
		\State \Return $ \Rule_S $
		\EndProcedure
	\end{algorithmic}
\end{algorithm}

@@

\begin{itemize}
	\item The first modification is with respect to the definition of coverage for decision sets. Unlike decision lists, we separate samples that are \textit{correctly covered} by $ (C_i, V_i) $ in each iteration. Given  a dataset $ \dataset = \{(\mathbf{x}^{(l)}, y^{(l)})\}$, the correctly covered samples of $ (C_i, V_i) $ is a dataset $ \dataset_{cc}  =  \{(\mathbf{x}^{(l)}, y^{(l)}) | \mathbf{x}^{(l)} \models C_i \wedge y^{(l)} = V_i \}_{l=1}^{|\dataset|} \subseteq \dataset $ of samples that satisfy $ C_i $ and have matching class-label as $ V_i $. 
	\item The second modification is related to the  training dataset considered in each iteration. Let $ \dataset $ denote \emph{the remaining training dataset} in the current iteration and $ V_i $ be the majority class in  $ \dataset $. Hence, $ V_i $ is the target class in the current iteration. Also, let $ \dataset_{cc} $ denotes the set of samples \emph{correctly covered in all previous iterations}. In Algorithm~\ref{interpretability_imli_algo:iterative_decision_sets_learning}, we learn the current rule $ C_i $ on the working dataset $ \dataset_{w} = \dataset  \cup \{(\mathbf{x}, \neg V_i) | (\mathbf{x}, y) \in \dataset_{cc}\} $. Thus, $ \dataset_{w} $ is constructed by joining the remaining training samples with correctly covered samples in $ \dataset_{cc} $ with complemented class label to the target class $ V_i $.  Thus, by explicitly labeling covered samples as class $ \neg V_i $, the new clause $ C_i $ learns to falsify already covered samples. This heuristic allows us to minimize the overlap of $ C_i $ compared to previously learned clauses $ \{C_j\}_{j=1}^{i-1} $.
\end{itemize}

### 2022-07-15 chapters/interpretability/relaxed-cnf/conclusion.tex
considering all minor comments

In this chapter, we proposed an efficient combinatorial framework called {\crr}, for learning relaxed-CNF  classification rules, that are  more expressive  than CNF/DNF rules.  {\crr}  uses a novel integration of mini-batch learning procedure with the MILP framework to learn sparse relaxed-CNF rules. Our experimental results demonstrate that {\crr} is able to learn relaxed-CNF  rules with higher accuracy and more concise representation than CNF rules. Moreover, the generated rules are sparser than decision lists in large datasets.

>>>>

In this chapter, we proposed an efficient combinatorial framework, called {\crr}, for learning relaxed-CNF  classification rules. Relaxed-CNF rules are  more expressive  than CNF/DNF rules.  {\crr}  uses a novel integration of mini-batch learning procedure with the MILP framework to learn sparse relaxed-CNF rules. Our experimental results demonstrate that {\crr} is able to learn relaxed-CNF  rules with higher accuracy and more concise representation than CNF rules. Moreover, the generated rules are sparser than decision lists in large datasets.
